---
title: "test"
author: "Yan Zichu"
date: "9/23/2020"
output: pdf_document
---
添加新的一列并且，这一列计算前面某几列
```{r}
companiesData$margin <- (companiesData$profit / companiesData$revenue) * 100
```
Using base R, write the R code to add an additional column to the gapminder dataset that gives the overall GDP of the country.
```{r}
gapminder$overallGDP <- (gapminder$pop * gapminder$gdpPercap)
```
Write R code to select only the columns "ref","company","company_location", "review_date" and "cocoa_percent" and "rating", and put them in a new data frame called choc_locations.

***************************************
Quiz 2
***************************************

```{r}
chocolate <- read.csv("chocolate_reduced.csv",stringsAsFactors = FALSE)
chocolate
```

```{r}
choc_locations <- chocolate[c("ref","company","company_location","review_date","cocoa_percent","rating")]
choc_locations
```
Write R code to extract the data for chocolate from companies in Switzerland, Belgium or New Zealand and put them in a new data frame called "choc_countries"


```{r}
choc_countries <- choc_locations[(choc_locations$company_location =="Switzerland"|
                                    choc_locations$company_location =="Belgium"|
                                    choc_locations$company_location =="New Zealand"), ]
choc_countries
```
Going back to the original dataset you read in from "chocolate_reduced.csv", write dplyr code to select the columns "cocoa_percent",  "counts_of_ingredients", "country_of_bean_origin" "rating", "lecithin" and "sugar" and put them in a new data frame called "choc_ingredients".
```{r}
library(dplyr)
library(tibble)
library(tidyr)
choc_ingredients<-select(chocolate, cocoa_percent, counts_of_ingredients,country_of_bean_origin,rating,lecithin,sugar)
choc_ingredients
```
```{r}
filter(choc_ingredients,cocoa_percent==100)
```
Write the R code to extract only those chocolate bars that have lecithin or sugar or both.

```{r}
filter(choc_ingredients,lecithin=="Yes"||sugar=="Yes")

```

Write the R code to convert the "counts_of_ingredients" column into a factor, with the levels of the factor being every integer from 2 to 6 in increasing order.

```{r}
mutate(choc_ingredients,level = cut(counts_of_ingredients, 5, labels=c("2", "3", "4", "5", "6")))

```
Write R code using ggplot2 to produce a boxplot of cocoa percent by counts of ingredients, with counts of ingredients on the x axis. Just give the R code to do this (you don't need to include the graphic image).
```{r}
library(ggplot2)
ggplot(choc_ingredients) +geom_boxplot(aes(x=counts_of_ingredients, y=cocoa_percent))

```
Using the data frame from the previous question, write R code to find the median rating for chocolate made with beans from each different country (i.e. the median rating for chocolate made with Australian beans, the median rating for chocolate made with Belgian beans, etc.).

Store the summary data in a new object called "choc_summary" and name the median rating column "median_rating".
```{r}
choc_summary <- group_by(choc_ingredients,country_of_bean_origin)
choc_summary <- summarise(choc_ingredients,country_of_bean_origin,median_rating=median(rating))
choc_summary<-arrange(choc_summary, desc(median_rating))

choc_summary


```


```{r}
ggplot(choc_ingredients, aes(x=cocoa_percent, color= lecithin,y =rating)) +geom_point() + 
  geom_point(position="jitter")
```



***************************************
Simulation
***************************************

```{r}
set.seed(2)                                                                    # 将随机数固定
sample(x=c(1,3,7),size=4,replace = TRUE,prob = c(0.1,0.3,0.6))
```
```{r}
outcome <- sample(x=c("heads","tails"),size=4,replace = TRUE,prob = c(0.5,0.5))
coin.flip <- data.frame(table(outcome))                                        # Quiz 扣分的地方！！！
coin.flip
```
```{r}
set.seed(1)                                                                   
N <- 100
r <- runif(N)
head(r)

outcomes <- rep(NA,N)
for(i in 1:N){
  if(r[i]<=0.5){
  outcomes[i] <- "head"
  }else{
  outcomes[i] <- "tail"
}
}
head(outcomes)
```
```{r}
outcomes <- ifelse(test = (r<=0.5),yes="heads",no="tails")
head(outcomes)

```
```{r}
dbinom(0 ,size = 1,prob = 0.7)
```
















